---
title: "Q2 Markdown"
output: 
  html_document:
    code_folding: hide  # Optionally hide code by default
---


```{r setup, include=FALSE}
# load packaged
rm(list = ls()) #clear env
library("fpp3") #load packages
library("tidyverse")
library("urca")
library("forecast")
library("rugarch")
library("knitr")

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(output = "asis")

# load data and get log returns 
v_5 = read.csv("data/V - 5 Years.csv") %>%  select(Date, Close) %>% 
  mutate(Date = date(Date)) %>% as_tsibble(index = Date) %>% 
  mutate(log_rtn = difference(log(Close))) %>% 
  filter(!is.na(log_rtn))

msft_5 = read.csv("data/MSFT - 5 Years.csv") %>%  select(Date, Close) %>% 
  mutate(Date = date(Date))%>% as_tsibble(index = Date) %>% 
  mutate(log_rtn = difference(log(Close))) %>% 
  filter(!is.na(log_rtn))

formulas_file = "q2_a.tex"
```


# Q 2 
The objective of this exercise is to obtain VaR on the basis of different as-
sumptions regarding the ARMA-GARCH process for the log returns. Use both
financial time series to answer the following question.

## a) 
Suppose that a given firm holds a long financial position of 10 Million Euros in stocks of the financial asset under analysis. Obtain the VaR for the next trading day and the next 5 trading days using:

```{r Q2 a setup}
value_stock = 10e6
cis = c(.90,.95,.99)
alphas = 1- cis

#### function for resutls ####

append_calc_string_list = function(string_list_loc, var_log_ret, q, m, s, alpha1, lag1, distr1 = "norm", value_stock1 = value_stock){
      string_list1 = c(string_list_loc, "\n",
                      
      #q,mu,sigma                
      paste0("\n\n",
             "Calculation for ", 100*(1-alpha1),"\\% Confidence and $T+", lag1, "$:\n\n",
             "\\indent\\indent $\\text{Quantile}_\\text{",distr1,"}(",alpha1,") = ",q,
             ",\\hat{\\mu_{T+",lag1,"|T}} = ",m,
             ", \\hat{\\sigma_{T+",lag1,"|T}} = ",s, "$\n"),
      
      #VaR log returns
      paste0("\\indent\\indent $VaR(\\log \\text{returns})_{",alpha1*100,"\\%, T + ",lag1,"} = ",m," + ",s,"\\times",q,
             " = ", round(var_log_ret, 4) ,"$\n"),
      
      #VaR Euro
      paste0("\\indent\\indent $VaR_{",alpha1*100,"\\%, T + ",lag1,"} = \\text{ Value in Euros } \\times VaR(\\log returns) = ",
             format(value_stock1, big.mark = " ", scientific = FALSE)," \\text{\\euro} \\times",
             round(var_log_ret, 4)," = ",round(value_stock1*var_log_ret, 2)," \\text{\\euro}$\\newline"))
  return(string_list1)     
}
      
      
getVaR_table = function(spec1, data1, lags = c(1:5), value_stock1 = value_stock, alphas1 = alphas, verbose = T,  formulas = T){
  # Function that predcits and forecasts using a rugarch::modelspec object and and tsibble data on log returns. it then gets the VaR for 2 specified lags in the future using a specified stock value and probability bounds
  #Parameters
  ## spec1: rugarch::modelspec object
  ## data1: log return data a tsibble object
  ## 2 lags: default c(1:5)
  ## value_stock1: default golbal var 'value_stock'
  ## alphas1: default golbal var 'alphas'
  ## verbose: should itermediary results be printed
  ## formulas: should latex formulas be written (to a seperate file)
  
  #fit model
  fit1 = rugarch::ugarchfit(data = data1, spec = spec1)
  #predict
  pred1 = rugarch::ugarchforecast(fit1,n.ahead = lags[length(lags)])
  
  #get pred mean and var
  mu1 = fitted(pred1)
  sig1 = sigma(pred1)
  
  distr1 = spec1@model$modeldesc$distribution
  
  if (distr1 == "norm"){
    quant_fun = qnorm
  } else if (distr1 == "std"){
    print("using student t quantilies")
    df_t = fit1@fit$matcoef["shape"," Estimate"]
    distr1 = paste0("std(", round(df_t), ")")
    quant_fun = function(ps){
      return(qt(p = ps, df = df_t))
    }
  } else {
    Warning("don't know this distribution, using nromal quantiles")
    quant_fun = qnorm
  }
  
  
  #get VaR
     #put lags and alphas into df
  max_neg_change_log_rtn <- data.frame(expand.grid(lags = lags, alphas1 = alphas1)) %>%
     #apply formual to everything
    mutate(result = (mu1[lags] + sig1[lags] * quant_fun(alphas1))) %>% 
    #change names
    mutate(lags = paste0(lags, "-Day"),
           alphas1 = paste0("VaR",(100*(alphas1)), "%")) %>% 
    #change format
    pivot_wider(names_from = alphas1, values_from = result) %>%
    column_to_rownames(var = "lags")
  
  max_neg_change_euro = value_stock1*max_neg_change_log_rtn

  if (verbose){
    # print(paste("Mean:", paste(mu1,  collapse = ",")))
    # print(paste("SD:",   paste(sig1, collapse = ",")))

    #print calculation resuslts as table
    ## quantiles
    print(paste("quantiles:", paste(round(quant_fun(alphas),2),collapse = ",")))
    
    # # % VaR
    # print(paste("         ",
    #             paste(names(max_neg_change_log_rtn),
    #                   collapse = " ")))
    # for (i in 1:nrow(max_neg_change_log_rtn)) {
    #   print(paste(rownames(max_neg_change_log_rtn)[i],
    #               paste(
    #                 paste0(
    #                   str_pad(round(max_neg_change_log_rtn,4)[i, ],
    #                       nchar("-0.0001"), "right"),
    #                   "% "),
    #                 collapse ="")
    #               
    #               ))
    # }  
    # 
    # # Euro VaR
    # 
    # print(paste("          ",
    #             paste(names(max_neg_change_euro),
    #                   collapse = "         ")))
    # for (i in 1:nrow(max_neg_change_euro)) {
    #   print(paste(rownames(max_neg_change_euro)[i],
    #               paste(
    #                 paste0(
    #                   str_pad(round(max_neg_change_euro,2)[i, ],
    #                           width = nchar("-400000.00"), "right"),
    #                 "Euros "), collapse ="")
    #               ))
    # } 
    
    ## both VaR
    print(paste("      mu     sig    ",
                paste(names(max_neg_change_log_rtn),
                      collapse = "   "),
                "    ",
                paste(names(max_neg_change_euro), collapse = "          ")))
    
    for (i in 1:nrow(max_neg_change_log_rtn)) {
      print(paste(
                  rownames(max_neg_change_log_rtn)[i],
                  str_pad(format(round(mu1[i],4), scientific = FALSE), 6, "right"),
                  str_pad(format(round(sig1[i], 4), scientific = FALSE), 6, "right"),
                  paste(
                    paste0(
                      str_pad(round(max_neg_change_log_rtn,4)[i, ],
                          nchar("-0.0001"), "right"),
                      "% "),
                    collapse =""),
                  paste(
                    paste0(
                      str_pad(round(max_neg_change_euro,2)[i, ],
                              width = nchar("-400000.00"), "right"),
                    "Euros "), collapse ="")
                ))
    }  

    
  }
  
  if (formulas){
    #print latex formulas
    string_list = c()
    
    for (lag in lags){
      for (alpha in alphas1){
        string_list = append_calc_string_list(string_list_loc = string_list,
                                              var_log_ret = mu1[lag] + sig1[lag]*quant_fun(alpha),
      q = round(quant_fun(alpha),2),
      m = format(round(mu1[lag],4), scientific = FALSE),
      s = format(round(sig1[lag],4), scientific = FALSE),
      alpha1 = alpha,
      lag1 = lag,
      distr1 = distr1)
      
      }
    }
    cat(string_list, file = formulas_file, append = TRUE, sep = "\n")
  }
  
  #concat df to return
  names(max_neg_change_log_rtn) = paste(names(max_neg_change_log_rtn), "(log rtn)")
  names(max_neg_change_euro) = paste(names(max_neg_change_euro), "(Euro)")
  desc = data.frame(mu_t = mu1[lags], sigma_t = sig1[lags])
  names(desc) = c("mu_t","sigma_t")
  return(cbind(desc,
               max_neg_change_log_rtn,
               max_neg_change_euro))
}

enum_str_list = function(string_list, sep = ", ", last = "  and "){
  #function that collapses a list of strings to a string.
  # "sep": seperator for all but the last element
  # "last": seperator for the last element
    return(
      paste0(
        c(
          paste0(string_list[1:(length(string_list)-1)], collapse = sep),
          string_list[length(string_list)]
          ),
        collapse = last
      )
    )
}

var_template_1_d = function(vaR,
                            stock_name,
                            day_num,
                            currency = "$\\text{\\euro}$",
                            alphas1 = alphas){
  day_text = switch(as.character(day_num),
    "1" = paste0("\n\nFollowing the modeling assumptions of this subtask, the available stock data and the given sum invested in the ", 
           stock_name,
           " stock, the loss incurring over one day will be smaller than "),
    "5" = "The loss after 5 days will be smaller than "
  )
  return(
    paste0(day_text,
           enum_str_list(
             paste0(
               format(
                 round(vaR,2),
                 nsmall = 2,
                 scientific = F
                 ),
               currency)
             ),
           "  with a confidence level of " ,
           enum_str_list(
             paste0(
               100*(1-alphas1),
               "\\%")
             ),
           ".\n"
           )
  )
}
  
```


### (i) 
#### the unconditional moments of an assumed normal distribution (historical  mean and standard deviation)

```{r start latex doc}
text_i = "\\section{Value at Risk (VaR) Analysis}
\\label{sec:q2_value_at_risk_analysis}
\\subsection{VaR calculations}
\\subsubsection{(i) Constant Model}

The only model that can be formed using just unconditional moments is the constant model $$R_t = \\epsilon_t, \\epsilon_t \\overset{\\text{w.n.}}{\\sim} N(\\mu, \\sigma^2),$$
with $\\hat \\mu = \\bar R_t, \\hat{\\sigma^2} = Var(R_t)$, for $t \\in [1,T]$. 
Even for a random walk model both the standard deviation of an observation from the preceding one and a last observation would be needed. Both are not given.


To calculate the VaR of each stock, the historical mean and standard deviation (sd) will be estimated and the respective z value of a standard distribution will be scaled by the sd and shifted by the mean. 
Since the likelihood of the z value can be read off from the normal distribution, the z value corresponding to the likelihood of interest can be chosen.


The scaled and shifted z value corresponds to a maximal log return under the given probability. The typically small log returns of stocks are proportional to percent changes in the stock value. Tue to this, maximal the monetary value of currency that is at risk can be calculated by multiplying the result of the previous calculation with the value of bought stock."

cat(text_i, file = formulas_file) 

text_i_5d = "The due to simplicity of the model, the 5-day VaR is the same as the 1-day VaR.\n"
```


``` {r i v 1 day}
#write to file
cat("\n\\paragraph{Visa} Caclulations.\\newline \\indent \n\n", file = formulas_file, append = TRUE)


## get mean and sd
v_m = v_5$log_rtn %>% mean()
v_sd = v_5$log_rtn %>% var() %>% sqrt()

## next day
p_v_1 = v_m + v_sd*qnorm(alphas)
v_a_i_max_1_day_neg_change = value_stock * p_v_1 

#print results
print(paste("mu: ", round(v_m,4), ", sd: ", round(v_sd, 4)))
print(paste("VaR(log rtn):", paste(round(p_v_1,4),collapse = ",")))
print(paste("VaR(Euro):", paste(round(v_a_i_max_1_day_neg_change, 2),collapse = ",")))


#print latex formulas
string_list = c()
    
lag = 1
for (alpha in alphas){
  string_list = append_calc_string_list(string_list_loc = string_list,
                                        var_log_ret = v_m[lag] + (v_sd[lag]*qnorm(alpha)),
                                        q = round(qnorm(alpha),2),
                                        m = format(round(v_m[lag],4), scientific = FALSE),
                                        s = format(round(v_sd[lag],4), scientific = FALSE),
                                        alpha1 = alpha,
                                        lag1 = lag)
  # string_list = c(string_list, "\n",
  # paste0("$\\text{Quantile}_\\text{norm}(",alpha,") = ",q,"
  #        \\hat{\\mu_{T+",lag,"|T}} = ",m,", \\hat{\\sigma_{T+",lag,"|T}} = ",s, "$\n"),
  # paste0("$VaR(\\log \\text{returns}) = ",m," + ",s,"\\times",q,
  #        " = ", round(var_log_ret, 4) ,"$\n"),
  # paste0("$VaR_{",alpha*100,"\\%} = \\text{ Value in Euros } \\times VaR(\\log returns) = ",
  #        format(value_stock, scientific = FALSE)," \\text{\\euro} \\times",
  #        round(var_log_ret, 4)," = ",round(value_stock*var_log_ret, 2)," \\text{\\euro}$"))
}

text_i_v_1d = var_template_1_d(v_a_i_max_1_day_neg_change, stock_name = "Visa", day_num = 1)

cat(c(string_list, text_i_v_1d,text_i_5d), file = formulas_file, append = TRUE, sep = "\n")
```


```{r i msft 1 day}
#write to file
cat("\n
\\paragraph{Microsoft} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

## get historical, unconditional mean and sd
msft_m = msft_5$log_rtn %>% mean()
msft_sd = msft_5$log_rtn %>% var() %>% sqrt()

## next day
p_msft_1 =  msft_m + (msft_sd* qnorm(alphas))
msft_a_i_max_1_day_neg_change = value_stock * p_msft_1 #assume proportionality to percents

#print results
print(paste("nomal quantiles:", paste(round(qnorm(alphas),2),collapse = ",")))
print(paste("mu: ", round(msft_m,4), ", sd: ", round(msft_sd, 4)))
print(paste("VaR(log rtn):", paste(round(p_msft_1,4),collapse = ",")))
print(paste("VaR(Euro):", paste(round(msft_a_i_max_1_day_neg_change, 2),collapse = ",")))

#print latex formulas
string_list = c()
    
lag = 1
for (alpha in alphas){
  string_list = append_calc_string_list(string_list_loc = string_list,
                                        var_log_ret = msft_m[lag] + msft_sd[lag]*qnorm(alpha),
                                        q = round(qnorm(alpha),2),
                                        m = format(round(msft_m[lag],4), scientific = FALSE),
                                        s = format(round(msft_sd[lag],4), scientific = FALSE),
                                        alpha1 = alpha,
                                        lag1 = lag)
  # string_list = c(string_list, "\n",
  # paste0("$\\text{Quantile}_\\text{norm}(",alpha,") = ",q,
  #        ",\\hat{\\mu_{T+",lag,"|T}} = ",m,", \\hat{\\sigma_{T+",lag,"|T}} = ",s, "$\n"),
  # paste0("$VaR(\\log \\text{returns}) = ",m," + ",s,"\\times",q,
  #        " = ", round(var_log_ret, 4) ,"$"),
  # paste0("$VaR_{",alpha*100,"\\%} = \\text{ Value in Euros } \\times VaR(\\log returns) = ",
  #        format(value_stock, big.mark = " ", scientific = FALSE)," \\text{\\euro} \\times",
  #        round(var_log_ret, 4)," = ",round(value_stock*var_log_ret, 2)," \\text{\\euro}$"))
}

text_i_msft_1d = var_template_1_d(msft_a_i_max_1_day_neg_change, stock_name = "Microsoft", day_num = 1)

cat(c(string_list, text_i_msft_1d,text_i_5d), file = formulas_file, append = TRUE, sep = "\n")
```

### (ii)
#### a GARCH(1,1) with only a constant in the conditional mean function and normal errors. Hint: for an ARMA(0,0) model, it is possible to show that ψi = 0 for every i.



```{r ii model}
text_ii = "The following GARCH model will be fitted 
$$\\begin{cases} R_t = \\mu_t + \\epsilon_t, \\epsilon_t = \\sigma_t  * z_t, z_t \\overset{i.i.d.}{\\sim} N(0,1) \\\\ \\sigma^2_t = \\alpha_0 + \\alpha_1 \\epsilon^2_{t-1} + \\beta_1 \\sigma^2_{t-1} \\\\ \\mu_t = \\mu_0 \\end{cases}$$

The model fitted in this subtask is more sophisticated than a simple model using only unconditional constants. 
The model will be specified and fitted to the historical data. The fitted model can then be used to make predictions about the stock returns in the future. The predicted mean and sd of the returns can be used to scale and shift a z value like in the previous subtask.
The use of this scaled and shifted z value in the further calculation of the VaR is identical to the previous, current and the later subtasks."

#write to file 
cat(c("\n\\subsubsection{(ii) AR(0)-GARCH(1,1)}\n",text_ii), file = formulas_file, append = TRUE)

#specify model 
a_ii_spec = rugarch::ugarchspec(variance.model=list(model="sGARCH", 
                                                 garchOrder=c(1,1)),
                             mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
                             distribution.model="norm")
```

```{r ii v}
#write to file
cat("\n\\paragraph{Visa} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
v_a_ii_table = getVaR_table(a_ii_spec, v_5$log_rtn)


#write conclusion
text_ii_v_1d  = var_template_1_d(v_a_ii_table[1,((ncol(v_a_ii_table)-2):(ncol(v_a_ii_table)))],
                                    stock_name = "Visa", day_num = 1)
text_ii_v_5d  = var_template_1_d(v_a_ii_table[5,((ncol(v_a_ii_table)-2):(ncol(v_a_ii_table)))],
                                    stock_name = "Visa", day_num = 5)

cat(c(text_ii_v_1d,text_ii_v_5d), file = formulas_file, append = TRUE, sep = "\n")
```


```{r ii msft}
#write to file
cat("\n\\paragraph{Microsoft} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
msft_a_ii_table = getVaR_table(a_ii_spec, msft_5$log_rtn)

#write conclusion
text_ii_msft_1d  = var_template_1_d(msft_a_ii_table[1,((ncol(msft_a_ii_table)-2):(ncol(msft_a_ii_table)))],
                                    stock_name = "Microsoft", day_num = 1)
text_ii_msft_5d  = var_template_1_d(msft_a_ii_table[5,((ncol(msft_a_ii_table)-2):(ncol(msft_a_ii_table)))],
                                    stock_name = "Microsoft", day_num = 5)

cat(c(text_ii_msft_1d,text_ii_msft_5d), file = formulas_file, append = TRUE, sep = "\n")

```

```{r ii final}
text_ii_final = "Based on this model, the VaR is smaller for investments into the Visa than for investiments into the Microsoft.

The all VaR obtained by this model are smaller than the VaR obtained by the simplistic model in the first subtask. That is due to the small time window, for which the time dependent variance and the VaR is computed. In this case, the time dependent variance is usually smaller than the historic variance."


cat(c(text_ii_final), file = formulas_file, append = TRUE, sep = "\n")

```

### (iii)
#### an AR(1)-GARCH(1,1) model with normal errors. Hint: for a stationary AR(1) model, it is possible to show that ψi = ϕ1i for every i.

```{r iii model}
#write to file
cat("\n\\subsubsection{(iii) AR(1)-GARCH(1,1)}
The following GARCH model will be fitted in this subtask
$$\\begin{cases} R_t = \\mu_t + \\epsilon_t, \\epsilon_t = \\sigma_t  * z_t, z_t \\overset{i.i.d.}{\\sim} N(0,1) \\\\ \\sigma^2_t = \\alpha_0 + \\alpha_1 \\epsilon^2_{t-1} + \\beta_1 \\sigma^2_{t-1} \\\\ \\mu_t = \\mu_0 + \\phi_i R_{t-1} \\end{cases}$$ 


here $\\mu_t$ is not only modeled as a constant but as a function of the past $\\mu_{t-1}$", file = formulas_file, append = TRUE)

#specify model
a_iii_spec <-  rugarch::ugarchspec(variance.model=list(model="sGARCH", 
                                                 garchOrder=c(1,1)),
                             mean.model=list(armaOrder=c(1,0),include.mean=TRUE),
                             distribution.model="norm")
```

```{r iii v}
#write to file
cat("\n\\paragraph{Visa} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
v_a_iii_table = getVaR_table(a_iii_spec, v_5$log_rtn)

#write to file
text_iii_v_1d  = var_template_1_d(v_a_iii_table[1,((ncol(v_a_iii_table)-2):(ncol(v_a_iii_table)))],
                                    stock_name = "Visa", day_num = 1)
text_iii_v_5d  = var_template_1_d(v_a_iii_table[5,((ncol(v_a_iii_table)-2):(ncol(v_a_iii_table)))],
                                    stock_name = "Visa", day_num = 5)


cat(c(text_iii_v_1d,text_iii_v_5d), file = formulas_file, append = TRUE, sep = "\n")
```


```{r iii msft}
#write to file
cat("\n\\paragraph{Microsoft} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
msft_a_iii_table = getVaR_table(a_iii_spec, msft_5$log_rtn)

#write to file
text_iii_msft_1d  = var_template_1_d(msft_a_ii_table[1,((ncol(msft_a_iii_table)-2):(ncol(msft_a_iii_table)))],
                                    stock_name = "Microsoft", day_num = 1)
text_iii_msft_5d  = var_template_1_d(msft_a_iii_table[5,((ncol(msft_a_iii_table)-2):(ncol(msft_a_iii_table)))],
                                    stock_name = "Microsoft", day_num = 5)

cat(c(text_iii_msft_1d,text_iii_msft_5d), file = formulas_file, append = TRUE, sep = "\n")
```

```{r iii final}
text_iii_final = "The estimated VaR don't really differ between the previous model which modeled the mean as constant, and the current one, which models the mean of the time series using an AR(1). 
The reason for this indifference could be, that the conditional mean of stock returns cannot really be predicted from preceding returns. It is already a good aproximation to model this conditional mean as constant. This is consitent with financial theory."

cat(c(text_iii_final), file = formulas_file, append = TRUE, sep = "\n")
```
### (iv) 
#### an AR(1)-GARCH(1,1) model with Student-t innovations 

```{r iv model}
#write to file
cat("\n\\subsubsection{(iv) AR(1)-GARCH(1,1) with t-Student innovations}
In this subtask, innovations of $R_t$ will not be modeled after a Normal distribution but a t-Student distibution (std) \n", file = formulas_file, append = TRUE)

#specify model
a_iv_spec <-  rugarch::ugarchspec(variance.model=list(model="sGARCH", 
                                                 garchOrder=c(1,1)),
                             mean.model=list(armaOrder=c(1,0),include.mean=TRUE),
                             distribution.model="std")
```

```{r iv v}
#write to file
cat("\n\\paragraph{Visa} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
v_a_iv_table = getVaR_table(a_iv_spec, v_5$log_rtn)

#write to file
text_iv_v_1d  = var_template_1_d(v_a_iv_table[1,((ncol(v_a_iv_table)-2):(ncol(v_a_iv_table)))],
                                    stock_name = "Visa", day_num = 1)
text_iv_v_5d  = var_template_1_d(v_a_iv_table[5,((ncol(v_a_iv_table)-2):(ncol(v_a_iv_table)))],
                                    stock_name = "Visa", day_num = 5)


cat(c(text_iv_v_1d,text_iv_v_5d), file = formulas_file, append = TRUE, sep = "\n")
```

```{r iv msft}
#write to file
cat("\n\\paragraph{Microsoft} Caclulations.\\newline \\indent \n", file = formulas_file, append = TRUE)

#get VaR
msft_a_iv_table = getVaR_table(a_iv_spec, msft_5$log_rtn)

#write to file
text_iv_msft_1d  = var_template_1_d(msft_a_iv_table[1,((ncol(msft_a_iv_table)-2):(ncol(msft_a_iv_table)))],
                                    stock_name = "Microsoft", day_num = 1)
text_iv_msft_5d  = var_template_1_d(msft_a_iv_table[5,((ncol(msft_a_iv_table)-2):(ncol(msft_a_iv_table)))],
                                    stock_name = "Microsoft", day_num = 5)


cat(c(text_iv_msft_1d,text_iv_msft_5d), file = formulas_file, append = TRUE, sep = "\n")
```


```{r iv final}
text_iv_final = "The VaR obtained by the current models using t-stundent innovations are much larger than the ones obtained by the similar models using normaly distributed innovations. This is caused by the shift of likelohood towards the tails in t-student distributions. This can reflect an increased likelihood of extreme events like negative price shocks."

cat(text_iv_final, file = formulas_file, append = T)
```


### (v) 
#### Another ARMA-GARCH specification choosen by the group. Here you should choose either a constant or an AR(1) conditional mean function. For the GARCH specification you may choose a GARCH(m,n) for any m and n with either normal or student-t innovations. You may also explore extensions of the standard GARCH model (GARCH-in-mean,T-GARCH, EGARCH,...). Justify your option

```{r v heading}
#write to file
cat("\n\\subsubsection{(v) Other ARMA-GARCH specification}\n", file = formulas_file, append = TRUE)
```

```{r v v}
#specify model
a_v_v_spec <-  rugarch::ugarchspec(variance.model=list(model="sGARCH", 
                                                 garchOrder=c(2,3)),
                             mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
                             distribution.model="std")

#write to file
cat(c("\n\\paragraph{Visa}\n",
      "An AR(0)-GARCH(2,3) model with normal innovations is used to obtain VaR values in this subtask. The model was chosen based on the problem statement and the Box-Jenkins-Methodology\n\n"),
    file = formulas_file, append = TRUE)

#get VaR
v_a_i_table = getVaR_table(a_v_v_spec, v_5$log_rtn)

#write to file
text_v_v_1d  = var_template_1_d(v_a_i_table[1,((ncol(v_a_i_table)-2):(ncol(v_a_i_table)))],
                                    stock_name = "Visa", day_num = 1)
text_v_v_5d  = var_template_1_d(v_a_i_table[5,((ncol(v_a_i_table)-2):(ncol(v_a_i_table)))],
                                    stock_name = "Visa", day_num = 5)

cat(c(text_v_v_1d,text_v_v_5d), file = formulas_file, append = TRUE, sep = "\n")
```

```{r v msft}
#specify model
a_v_msft_spec <-  rugarch::ugarchspec(variance.model=list(model="sGARCH", 
                                                 garchOrder=c(2,2)),
                             mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
                             distribution.model="std")

#write to file
cat(c("\n\\paragraph{Microsoft}\n",
    "An AR(0)-GARCH(2,2) model with normal innovations is used to obtain VaR values in this subtask. The model was chosen based on the problem statement and the Box-Jenkins-Methodology\n\n"),
    file = formulas_file, append = TRUE)

#get VaR
msft_a_i_table = getVaR_table(a_v_msft_spec, msft_5$log_rtn)

#write to file
text_v_msft_1d  = var_template_1_d(msft_a_i_table[1,((ncol(msft_a_i_table)-2):(ncol(msft_a_i_table)))],
                                    stock_name = "Microsoft", day_num = 1)
text_v_msft_5d  = var_template_1_d(msft_a_i_table[5,((ncol(msft_a_i_table)-2):(ncol(msft_a_i_table)))],
                                    stock_name = "Microsoft", day_num = 5)


cat(c(text_v_msft_1d,text_v_msft_5d), file = formulas_file, append = TRUE, sep = "\n")
```

# Functional Code
```{r knit formulas file}
rmarkdown::render(input = formulas_file, output_format = "html_document")
```

